%! TeX root = ../*.tex
\renewcommand{\pmk}{Bunks\_1995\_Geophy\_Multiscale waveform inversion}
\renewcommand{\prf}{FWI/\pmk.pdf}
\renewcommand{\pti}{Multiscale seismic waveform inversion}
\renewcommand{\pay}{Carey Bunks, Fatimetou M. Saleck, S. Zaleski,
  and G. Chavent, 1995}
\renewcommand{\pjo}{Geophysics}
\renewcommand{\pda}{2019/6/16 Sun.}

\section{\pinfo}

\subsection{Intruoduction}
At long scales there are fewer local minima, and those that remain
are further apart from each other.

\begin{enumerate}[\hspace{10mm}*]
  \item Linearized waveform inversion: Berkhout, 1984; Devaney, 1984;
    Esmersoy, 1986; Levy and Esmersoy, 1988; Tarantola, 1984.
  \item \sline
  \item Full nonlinear waveform velocity inversion: Mora, 1987;
    Pica \etal, 1990; Tarantola, 1986 \& 1988.
  \item \sline
  \item Reflection tomography method: Bishop \etal, 1985; Bording \etal, 1987.
  \item \sline
  \item Reduce the number of local minima in the objective function by
    introducing geometrically coherent constrains on the observed seismic
    data after depth migration: Al-Yahya, 1987; Bunks, 1991 \& 1992;
    Chavent and Jacewitz, 1990; Symes and Carazzone, 1991; van Trier, 1990.
  \item \sline
  \item Linear search: Luenberger, 1969.
  \item Lagrange multiplier technique: Alex\'eev \etal, 1982; Hildebrand, 1965;
    Lanczos, 1962; Luenberger, 1969.
  \item The classic second-order finite-difference scheme:
    Bamberger \etal, 1980; Kelly \etal, 1976.
  \item The multigrid method: Brandt, 1977; Briggs, 1987;
    Press and Teukolsky, 1991.
  \item The Jacobi and the Gauss-Seidel operators: Briggs, 1987.
  \item A quasi-Newton algorithm: Luenberger, 1989.
  \item Simulated annealing: Geman and Geman, 1984; Kirkpatrick \etal, 1983;
    Marroquin, 1985; Metropolis \etal, 1953.
  \item FIR Hamming windowed low-pass filter: Oppenheim and Schafer, 1975;
    Rabiner and Gold, 1975.
\end{enumerate}

Linearized inversion is justified when the initial velocity model is
in the neighborhood of the global minimum of the the objective function.

The observable wave vector components of the velocity field are bounded
above by the highest frequency in the source and
below by the effective opening of the seismic array.

The main theoretical difficulty for nonlinear seismic inversion is
the presence of numerous local minima in the objective function.

\subsection{The multigrid method}
The implementation of this algorithm requires three elements.
The restriction operator (go down to longer scale from shorter scale):
a leaky (large transition band) low-pass filter;
The relaxation operator (solve the longer scale): a quasi-Newton algorithm;
The injection operator (back up to shorter scale from longer scale):
the adjoint of a nine-point nearest-neighbor smoothing filter.

% vim:sw=2:wrap
